### 2015-02-27 icfp15-servo/browsers.tex
Edits from Simon

A URL identifies a resource to load.
This resource usually consists of HTML, which is then parsed and typically turned into a Document Object
Model (DOM) tree.
From a programming languages standpoint, there are several interesting aspects of the parser design
for HTML.
First, though the specification allows the browser to abort on a parse error\footnote{\url{https://html.spec.whatwg.org/multipage/syntax.html#parsing}},
in practice browsers follow the recovery algorithms described in that specification and make a best attempt to
load any page given to them.
Second, due to the presence of the \lstinline[language=HTML]{<script>} tag, the token stream can be modified
during operation.
For example, the below example that injects a close tag for the script block works in all modern browsers.
\begin{lstlisting}[language=HTML]
<html>
  <script>
    document.write (``</script>
  <li>this shows up just fine, thanks</li>
</html>
\end{lstlisting}
Finally, since resource loading is such a large factor in the latency of loading many webpages (particularly on mobile),
all modern parsers also perform token stream scanning and prefectch of resources likely to be required~\cite{browsers-slow-smartphones}.

>>>>

A URL identifies a resource to load.
This resource usually consists of HTML, which is then parsed and typically turned into a Document Object
Model (DOM) tree.
From a programming languages standpoint, there are several interesting aspects of the parser design
for HTML.
First, though the specification allows the browser to abort on a parse error\footnote{\url{https://html.spec.whatwg.org/multipage/syntax.html#parsing}},
in practice browsers follow the recovery algorithms described in that specification precisely so that
even ill-formed HTML will be handled in an interoperable way across all browsers..
Second, due to the presence of the \lstinline[language=HTML]{<script>} tag, the token stream can be modified
during operation.
For example, the below example that injects a close tag for the script block works in all modern browsers.
\begin{lstlisting}[language=HTML]
<html>
  <script>
  document.write(``<h'');
  </script>1>
  This is a h1 title

  <script>
  document.write(``<!-'');
  </script>-
  This is commented
</html>
\end{lstlisting}
This requirement causes parsing to pause while JavaScript code is running.
But, since resource loading is such a large factor in the latency of loading many webpages (particularly on mobile),
all modern parsers also perform token stream scanning and prefetch of resources likely to be required~\cite{browsers-slow-smartphones}.
